A Triple Origin for Twin Blue Stragglers in Compact Binaries

### 2018-12-04 apjl.tex
Add files via upload

Consider a hierarchical triple system with component masses m$_{\rm 1}$ and m$_{\rm 2}$ for the inner binary, and mass m$_{\rm 3}$ for the outer tertiary companion.  The inner and outer binary orbital semi-major axes are denoted a$_{\rm in}$ and a$_{\rm out}$, respectively.  We assume circular orbits for both the inner and outer orbits, and co-planar triples only since this is what has been observed for low-mass tertiaries \citep[e.g.][]{moe18,tobin18}.

@@

Figure~\ref{fig:fig1} shows the parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial component masses of m$_{\rm 1} =$ 1.1 M$_{\rm \odot}$ and m$_{\rm 2} =$ 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3} =$ 1.4 M$_{\rm \odot}$ for the outer tertiary.  We compare the circularization radius to the semi-major axis of the inner binary, for which we require a$_{\rm c} >$ a$_{\rm in}$, after folding in all constraints from the requirements for dynamical stability, an outer tertiary that is Roche lobe-filling, and a dynamically hard outer tertiary orbit.  Note that the range of plotted orbital periods P$_{\rm in}$ corresponding to a contact state for the inner binary, assuming R$_{\rm 1} =$ R$_{\rm 2} =$ 1 R$_{\odot}$, lies outside the range of plotted values for P$_{\rm in}$, since it does not contribute significantly to constraining the outer orbital properties of a hypothesized outer WD tertiary.  The thick horizontal solid red line shows the allowed range of outer semi-major axes, after folding in all of the aforementioned criteria.  As is clear, this makes a relatively narrow prediction for the allowed ranges of outer tertiary orbits, namely 2.2 $\times$ 10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, for our assumed final donor mass.

\begin{figure}[ht!]
\plotone{fig1.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane allowed for the hypothetical outer tertiary orbit of Binary 7782.  The solid diagonal black lines show the period corresponding to the circularization radius a$_{\rm c}$ for the mass transfer stream coming from the outer tertiary, for different values of the mass ratio, namely q $=$ 0.1, 0.7, 1, 2 and 10.  We assume initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ is computed for the outer tertiary according to our assumed mass ratio (with our fiducial case corresponding to q $=$ 0.7.  We assume completely conservative mass transfer for this exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal black line shows a rough criterion for dynamical stability in the triple, approximately following \citet{mardling99} (i.e., a$_{\rm in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical stability in equal-mass co-planar triples).  The vertical solid red lines show the outer orbital periods corresponding to the hard-soft boundary assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km s$^{-1}$.  The vertical dashed black lines show the maximum outer orbital period P$_{\rm out}$ for which the outer tertiary companion is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$ 200 R$_{\odot}$ (CHANGE, POSSIBLY USING MESA CALCULATIONS?!).  The horizontal dashed red line shows the observed orbital period for Binary 7782, using its observed orbital period and our assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line shows the parameter space for P$_{\rm out}$ allowed after considering all of the aforementioned criteria.
\label{fig:fig1}}
\end{figure}

@@

We use smoothed-particle hydrodynamics (SPH) simulations to compute the time evolution of the mass transfer process.  We adopt the same initial particle masses as in Figure~\ref{fig:fig1}, namely initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ is computed for the outer tertiary according to our assumed mass ratio, with our fiducial case corresponding to q $=$ 0.7.  The initial orbital eccentricities are set to zero.  The inner and outer binary orbital semi-major axes are set to 0.1 AU and 1 AU, respectively (CORRECT).

@@

\item The BSs in the inner binary should have roughly equal masses, independent of their initial masses.  This is because it is the lowest mass object that typically accretes the fastest, since its orbital velocity and distance relative to the circumbinary disk is typically the lowest \citep[e.g.][]{haiman09,farris15,rafikov16,kelley17}.  This quickly brings the mass ratio toward unity.  This predicts that the initially lower mass MS star should accrete the most, and should hence be polluted more significantly by any accreted material.  This could be observable in the surface layers of a radiative star.  If the donor is an RGB star, the accretor will be enriched in mostly carbon, oxygen and helium.  If the accretor is an AGB star, it will be enriched in mostly s-process elements.

>>>>

Consider a hierarchical triple system with component masses m$_{\rm 1}$ and m$_{\rm 2}$ for the inner binary, and mass m$_{\rm 3}$ for the outer tertiary companion.  The inner and outer binary orbital semi-major axes are denoted a$_{\rm in}$ and a$_{\rm out}$, respectively.  We assume circular orbits for both the inner and outer orbits, and co-planar triples only since this is what has been observed for low-mass tertiaries \citep[e.g.][]{moe18,tobin18}.  This initial configuration for our assumed formation scenario for Binary 7782, described below, is depicted in Figure~\ref{fig:fig1}.

\begin{figure}[ht!]
\plotone{fig1.eps}
\caption{Cartoon depiction of our proposed scenario for the formation of Binary 7782, specifically mass transfer from an evolved outer tertiary companion on to a compact inner binary via a circumbinary disk.  The outer tertiary component has mass m$_{\rm 3}$, whereas the inner binary components have masses m$_{\rm 1}$ and m$_{\rm 2}$.  The inner and outer orbital separations are denoted by, respectively, a$_{\rm in}$ and a$_{\rm out}$.  The circularization radius of the accretion stream is denoted a$_{\rm c}$, as calculated via Equation~\ref{eqn:ac}, and marks the mean separation of the circumbinary disk.
\label{fig:fig1}}
\end{figure}

@@

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial component masses of m$_{\rm 1} =$ 1.1 M$_{\rm \odot}$ and m$_{\rm 2} =$ 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3} =$ 1.4 M$_{\rm \odot}$ for the outer tertiary.  We compare the circularization radius to the semi-major axis of the inner binary, for which we require a$_{\rm c} >$ a$_{\rm in}$, after folding in all constraints from the requirements for dynamical stability, an outer tertiary that is Roche lobe-filling, and a dynamically hard outer tertiary orbit.  Note that the range of plotted orbital periods P$_{\rm in}$ corresponding to a contact state for the inner binary, assuming R$_{\rm 1} =$ R$_{\rm 2} =$ 1 R$_{\odot}$, lies outside the range of plotted values for P$_{\rm in}$, since it does not contribute significantly to constraining the outer orbital properties of a hypothesized outer WD tertiary.  The thick horizontal solid red line shows the allowed range of outer semi-major axes, after folding in all of the aforementioned criteria.  As is clear, this makes a relatively narrow prediction for the allowed ranges of outer tertiary orbits, namely 2.2 $\times$ 10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, for our assumed final donor mass.

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane allowed for the hypothetical outer tertiary orbit of Binary 7782.  The solid diagonal black lines show the period corresponding to the circularization radius a$_{\rm c}$ for the mass transfer stream coming from the outer tertiary, for different values of the mass ratio, namely q $=$ 0.1, 0.7, 1, 2 and 10.  We assume initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ is computed for the outer tertiary according to our assumed mass ratio (with our fiducial case corresponding to q $=$ 0.7.  We assume completely conservative mass transfer for this exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal black line shows a rough criterion for dynamical stability in the triple, approximately following \citet{mardling99} (i.e., a$_{\rm in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical stability in equal-mass co-planar triples).  The vertical solid red lines show the outer orbital periods corresponding to the hard-soft boundary assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km s$^{-1}$.  The vertical dashed black lines show the maximum outer orbital period P$_{\rm out}$ for which the outer tertiary companion is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$ 200 R$_{\odot}$ (CHANGE, POSSIBLY USING MESA CALCULATIONS?!).  The horizontal dashed red line shows the observed orbital period for Binary 7782, using its observed orbital period and our assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line shows the parameter space for P$_{\rm out}$ allowed after considering all of the aforementioned criteria.
\label{fig:fig2}}
\end{figure}

@@

We use smoothed-particle hydrodynamics (SPH) simulations to compute the time evolution of the mass transfer process.  We adopt the same initial particle masses as in Figure~\ref{fig:fig2}, namely initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ is computed for the outer tertiary according to our assumed mass ratio, with our fiducial case corresponding to q $=$ 0.7.  The initial orbital eccentricities are set to zero.  The inner and outer binary orbital semi-major axes are set to 0.1 AU and 1 AU, respectively (CORRECT).

@@

\item The BSs in the inner binary should have roughly equal masses, independent of their initial masses.  This is because it is the lowest mass object that typically accretes the fastest, since its orbital velocity and distance relative to the circumbinary disk is typically the lowest \citep[e.g.][]{haiman09,farris15,rafikov16,kelley17}.  This quickly brings the mass ratio toward unity.  This predicts that the initially lower mass MS star should accrete the most, and should hence be polluted more significantly by any accreted material.  This could be observable in the surface layers of a radiative star.  If the donor is an RGB star, the accretor will be enriched in mostly carbon, oxygen and helium.  If the donor is an AGB star, it will be enriched in mostly s-process elements.
